Centralized Programming for Distributed Control Planes

### 2016-01-25 paper/motivation.tex
minor changes to motivation

If configuring policies for a fully functional network is difficult, doing so in a way that ensures policy compliance in the face of failures can be almost impossible. Consider the data center network in Figure~\ref{fig:example2} with routers organized as a fat tree and running BGP.\footnote{To scale and to simplify policy implementation, data center networks increasingly use BGP internally, with a private AS number per router~\cite{bgp-in-dc-rfc}.} The network has two clusters, one that hosts services that should be reachable globally and one that hosts that should be accessible only internally. This policy is enabled by using non-overlapping address space in the two clusters and ensuring that only the address space for the global services is announced externally. In addition, to reduce the number of prefixes that are announced externally, the global space is aggregated into one larger, less-specific prefix $P_G$. The semantics of aggregation in BGP is that the aggregate prefix is announced as long as the router has a path to even one sub-prefix.

The operator may decide that a simple way to implement the policy is to have X and Y: $i)$ not export externally what they hear from routers G and H since these routers border the cluster with local services; and $ii)$ announce what they hear from routers C and D and aggregate to $P_G$ if an announcement is subset of $P_G$. The appeal of this implementation is that X and Y do not need to be made aware of which prefixes are global versus local and IP address assignment can be done independently (e.g., new prefixes can be added to local services without updating router configurations).

However, this implementation is incorrect because it does not have the right behavior in the face of failures. Suppose links X--G and X--H fail. Then, X will hear announcements for $P_{l*}$ from C and D, having traversed from G and H to Y to C and D. Per policy implementation, X will start "leaking" these prefixes externally. Depending on the rationale for local services, this leak could impact security (e.g., if the services were sensitive) or availability (e.g., if the $P_{l*}$ prefixes are used for other services outside of the data center). This problem does not manifest without failures because then X has and prefers paths to $P_{l*}$ through G and H since they are shorter. A similar problem will happen if links Y--G and Y--H fail.
%\footnote{A different problem occurs when links X--C and X--D (or Y--C and Y--D) fail. X (Y) may stop announcing the global prefixes because they would be heard through G and H.}
Link failures in data centers are frequent and it is not uncommon to have multiple failed links at a given time~\cite{dc-failure-study}.

To avoid this problem, the operator may decide to disallow paths with "valleys," i.e., those that go up, down, and back up again. This safeguard can be implemented by $X$ and $Y$ rejecting paths through the other. However, that creates a different problem in the face of failures---an aggregation-induced blackhole~\cite{xx}. Suppose links D--A and X--C fail. Now, X will hear announcements for $P_{g2}$ from D and will thus announce the $P_G$ externally. This announcement will bring to X traffic for $P_{g1}$ as well, but because of valley-free filtering, X does not have a valid route for $P_{g1}$ and will thus drop all traffic to it.

Thus, we see that devising a configuration that ensures policy compliance in the face of failures is complex and error-prone. \sysname helps operators by implementing their high-level policy specification in a way that guarantees compliance under all failures if that is possible. (Otherwise, it generates a compile-time error that informs them that the specification cannot be met.) For aggregation, it will also provide a lower bound to operators on the number of failures under which aggregation will not result in blackholes.

>>>>

 While configuring policies for a fully functional network is difficult, ensuring policy compliance in the face of failures can be almost impossible. Consider the data center network in Figure~\ref{fig:example2} with routers organized as a fat tree and running BGP.\footnote{For scale and policy flexibility, data center networks increasingly use BGP internally, with a private AS number per router~\cite{bgp-in-dc-rfc}.} The network has two clusters, one with services that should be reachable globally and one with services that should be accessible only internally. This policy is enabled by using non-overlapping address space in the two clusters and ensuring that only the address space for the global services is announced externally. Further, to reduce the number of prefixes that are announced externally, the global space is aggregated into a less-specific prefix $P_G$. The semantics of aggregation is that the aggregate prefix is announced as long as the router has a path to at least one sub-prefix.

The operator may decide that a simple way to implement the policy is to have X and Y: $i)$ not export externally what they hear from G and H, routers that belong to the local services cluster; and $ii)$ announce what they hear from routers C and D and aggregate to $P_G$ if an announcement is subset of $P_G$. This implementation is appealing because X and Y do not need to be made aware of which prefixes are global versus local and IP address assignment can be done independently (e.g., new prefixes can be assigned to local services without updating router configurations).

However, this implementation does not have the right behavior in the face of failures. Suppose links X--G and X--H fail. Then, X will hear announcements for $P_{l*}$ from C and D, having traversed from G and H to Y to C and D. Per policy implementation, X will start "leaking" these prefixes externally. Depending on the rationale for local services, this leak could impact security (e.g., if the services were sensitive) or availability (e.g., if the $P_{l*}$ prefixes are used for other services outside of the data center). This problem does not manifest without failures because then X has and prefers paths to $P_{l*}$ through G and H since they are shorter. A similar problem will happen if links Y--G and Y--H fail.
%\footnote{A different problem occurs when links X--C and X--D (or Y--C and Y--D) fail. X (Y) may stop announcing the global prefixes because they would be heard through G and H.}
Link failures in data centers are frequent and it is not uncommon to have multiple failed links at a given time~\cite{dc-failure-study}.

To avoid this problem, the operator may decide to disallow "valleys" paths, i.e., those that go up, down, and back up again. This guard can be implemented by $X$ and $Y$ rejecting paths through the other. But that creates a different problem in the face of failures---an aggregation-induced blackhole~\cite{xx}. If links D--A and X--C fail, X will hear announcements for $P_{g2}$ from D and will thus announce the $P_G$ externally. This announcement will bring to X traffic for $P_{g1}$ as well, but because of valley-free filtering, X does not have a valid route for $P_{g1}$ and will thus drop all traffic to it.

Thus, we see that devising a configuration that ensures policy compliance in the face of failures is complex and error-prone. \sysname lets operators implement their high-level policy specification in a way that guarantees compliance under all failures (if that is possible; otherwise, it generates a compile-time error). For aggregation, it will also provide a lower bound to operators on the number of failures under which aggregation will not result in blackholes.
